\section{Introduction}
In essential programming doctrine, it is considered ``best practice'' to write code that
is concise, understandable, and simple. This paradigm is great for sharing ideas
with other programmers, but doesn't hold up when we consider some business and
security realities of the world. For one, having unknown people understand
your code is not always part of the goal; consider a company that produces
high-performance code where much of their business comes from the fact that
other companies cannot reproduce their algorithm. If they sell or
release their software, they would want it to be done in a way that others still
cannot understand or reproduce its innerworkings.
Similarly, if code is easy to understand and contains any vulnerabilities,
those vulnerabilities will be easier to spot and exploit by unknown, possibly
malicious actors. Even for relatively secure code, understandability can reveal
too much of its innerworkings to the public and result in an attack. On the 
other side of legality, those writing malicious code want their code to be 
as opaque as possible so that the professionals working against them are not
able to mitigate the impact of their code or even detect its malicious intent
as effectively.
\par So, many kinds of programmers want to write code that is understandable
while they write it, but for several different reasons may want to release or
use a version of that code that is obfuscated so that it is not
able to be reverse engineered as easily. How can we reconcile
these conflicting goals? The primary method is {\itshape software obfuscation}.
Obfuscation is a process performed on code that transforms it into other code
which performs essentially the same operations, but is much more difficult to
understand or emulate. There are several technical ways to do this, obfuscating
the control flow of the program or using clever hardware/system-specific tricks
to dissuade reverse engineers. But these won't always help us hide special values
and computations that might also be sensitive information--for that, we'll have to
look to linear algebra.

In this paper, we discuss two techniques of code obfuscation. We first
introduce bitwise operators and mixed boolean arithmetic to obfuscate
expressions. We then introduce data types and explain how they and
their operations can be represented as matrices. In both sections, we provide a
simple function that we attempt to obfuscate using the aforementioned
techniques in order to contrast them. We also provide throughout the paper
Python code samples that are equivalent to our mathematical functions, in order
to show how actual code might be obfuscated using these methods.
